---
title: "Chapter 3 Part I Linear Regression (Problems_8)"
author: "Kartheek Raj"
date: "12/21/2019"
output:
  pdf_document: default
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Problem 8: This question involves the use of simple linear regression on the Auto data set
Required Packages :ISLR

Answer

Loading the **Auto** data from ISLR package
```{r Auto}
require(ISLR)
library(ISLR)
auto<-data.frame(Auto)
head(auto)
```
### (a) Use the lm() function to perform a simple linear regression with mpg as the response and horsepower as the predictor. Use the summary() function to print the results. Comment on the output.

Answer

```{r Linearmodel}
autolm<-lm(mpg~horsepower,data=auto)
summary(autolm)
```
#### (i) Is there a relationship between the predictor and the response? 

Answer: There is relationship between the predictor and the response as evidence poised to P-value.

#### (ii) How strong is the relationship between the predictor and the response?

Answer: Adjusted R-squared value is 0.60 means 60% can explain variance in **mpg** by **horsepower**.

#### (iii) Is the relationship between the predictor and the response positive or negative? 

Answer: Negative correlation indicated by horsepower coefficient is -0.157845 i.e m=negative slope.

#### (iv) What is the predicted mpg associated with a horsepower of 98? What are the associated 95% conï¬dence and prediction intervals? 

Answer

```{r cofidenceintervals}
predict(autolm,data.frame(horsepower=98),interval="confidence")

```
```{r predictions}
predict(autolm,data.frame(horsepower=98),interval="prediction")

```
### (b) Plot the response and the predictor. Use the abline() function to display the least squares regression line.

Answer

```{r plotpredictions}
plot(auto$horsepower,auto$mpg,main="Plot using adbline() display the least squares regression line",xlab="Horsepower",ylab = "mpg",col="darkorange")
abline(autolm,col="darkblue")

```
### (c) Use the plot() function to produce diagnostic plots of the least squares regression fit. Comment on any problems you see with the fit.

Answer

```{r diagnosticsplot}
plot(autolm,col="darkgreen")

```

