---
title: "Chapter 10 Problems 7"
author: "Kartheek Raj"
date: "12/26/2019"
output:
  html_document: default
  pdf_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 7. In the chapter, we mentioned the use of correlation-based distance and Euclidean distance as dissimilarity measures for hierarchical clustering. It turns out that these two measures are almost equivalent: if each observation has been centered to have mean zero and standard deviation one, and if we let rij denote the correlation between the ith and jth observations, then the quantity 1−rij is proportional to the squared Euclidean distance between the ith and jth observations. On the USArrests data, show that this proportionality holds. 

Required packages : ISLR

Answer

Data pulling from ISLR Library 
```{r USArrests}
require(ISLR)
library(ISLR)
attach(USArrests)
```

Proportionality test Euclidean distance vs Correlation distance as mentioned in the question.
```{r prop holds}

phtest<-function(x){
  sx<-scale(x)
  distdat<-dist(sx)
  sqrdistdat<-(distdat)^2
  tsx<-t(sx)
  cor11<-cor(tsx)
  cordata<-as.dist(1-cor11)
  pro<-(sqrdistdat/cordata)
  print(summary(pro))
  par(mfrow=c(1,2),bg="lightyellow")
  plot(pro)
  plot(pro,ylim = c(0,200))
  summary(pro)
  abline(h=3,col="red",lty="dashed")
  abline(h=14,col="green",lty="dashed")
}
paste0(phtest(USArrests))
paste0(phtest(USArrests[1:20,]))
paste0(phtest(USArrests[1:30,]))

```

Insight 1 : As we plotted whole and samples its followed the pattern most of theobervation  falls under green and blue dotted lines with some  noise showing evidence proportionality holds.


