---
title: "Chapter 3 Part I Linear Regression (Problem  12)"
author: "kartheek raj mulasa"
date: "12/21/2019"
output:
  pdf_document: 
   extra_dependencies:
      hyperref: ["unicode=true", "breaklinks=true","bbm", "threeparttable"]
      lmodern: null
  word_document: default
  html_document: default
  
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)


```

## Problem 12: This problem involves simple linear regression without an intercept.

### (a) Recall that the coeﬃcient estimate ˆ β for the linear regression of Y onto X without an intercept is given by (3.38). Under what circumstance is the coeﬃcient estimate for the regression of X onto Y the same as the coeﬃcient estimate for the regression of Y onto X? 

Answer

In generic linear equation y=mx+c,C=0 as intercept is zero then equation y=mx viceversa x onto y gives equation x=my both linear equations have the same slope occurs when y=x.

Given the linear regression equation in (3.38) we can say denmoniator term must be equal as numerator same for both events \beta=\sum_{i}_x_{i}y_{i}

$$ 
\sum_{{i^`}=1}^n x^2_{i^`} = \sum_{{i^`}=1}^n y^2_{i^`}
$$

### (b) Generate an example in R with n = 100 observations in which the coeﬃcient estimate for the regression of X onto Y is diﬀerent from the coeﬃcient estimate for the regression of Y onto X.

Answer 

creating the obersvations
```{r 100obesrvations}
set.seed(1)
x<-sample(1:1000,100,replace = T)
y<-x*2+sample(1:1000,100,replace = T)
length(x)
length(y)
```
Building linear models x onto y and y onto x.
```{r linearmodel1}
slm<-lm(y~x+0)
slm1<-lm(x~y+0)
summary(slm)
summary(slm1)
```
Generated linear models yields different coefficents

### (c) Generate an example in R with n = 100 observations in which the coeﬃcient estimate for the regression of X onto Y is the same as the coeﬃcient estimate for the regression of Y onto X. 

Answer 

Generating the 100 obervations
```{r 100obervations 1}

x<-100:200
x<-x^2
y<-200:100
y<-y^2
```
Building linear models x onto y and y onto x.
```{r linearmodels 2}
slm2<-lm(x~y+0)
slm3<-lm(y~x+0)
summary(slm2)
summary(slm3)
```
Both models yield same coefficent.